Information about the vertical state of the atmosphere is of high value
in many areas of meteorology. Vertical profiles of temperature and humidity are
used to determine atmospheric stability and the state of the boundary layer,
have a history of use in day-to-day weather forecasting and are routinely
assimilated into numerical weather prediction models. Currently, in-situ
balloon soundings (radiosoundings), aircraft measurements and remote sensing
data from radar or space-bourne instruments are main sources of information
about the vertical state of the atmosphere
\cite[alternative=authoryears,left={(e.g. }][Brousseau2014]. In recent years,
ground-based radiometers have been established as an additional source of
vertical information about the atmosphere. These passive remote sensing
instruments have been used for decades in research and are increasingly tested
for operational use. Additionally projects and user networks specialized on
microwave radiometers like the Atmospheric Radiation Measurement Program
\cite[authoryears][Cadeddu2013] and MWRnet
(\hyphenatedurl{http://cetemps.aquila.infn.it/mwrnet/}\footnote{last accessed
2016-08-04}) have emerged.


\startsection[title={Properties and Applications of Microwave Radiometers}]

    Measurements from a ground-based microwave radiometers (MWR) can be used to
    determine vertical profiles of temperature, humidity and cloud liquid water
    content \cite[authoryears][Westwater2004] depending on the choice of
    channels which are observed by the instrument. In practice it has been
    found that temperature retrievals generally achieve higher accuracy than
    humidity retrievals, partially due to the microwave absorption properties
    of the atmosphere \cite[authoryears][Lohnert2004,Xu2014]. Radiative
    transfer in the microwave region is dominated by absorption and emission of
    oxygen, water vapor and liquid water droplets. Cloud absorption weak enough
    to allow penetration of microwave radiation therefore it is possible to
    perform radiometer measurements even during clouds, giving MWRs an
    advantage over infrared radiometers which are used for similar purposes
    \cite[authoryears][Westwater2004,Kadygrov2015]. The effect of ice particles
    on microwave radiation is negligible, but scattering caused by
    precipitation is too strong to be neglected and retrievals lose accuracy
    (REF, mention that elevation scanning might help?).
    
    A radiometer is a passive remote sensing instrument. It can therefore
    be used in settlements \cite[authoryears][Kadygrov2015] and generally
    requires less maintenence and upkeep cost than an active remote sensor
    \cite[authoryears][Guldner2001]. In the long term it is also much
    cheaper than periodically launching balloon-bourne in-situ instruments.
    A disadvantage of being a passive instrument is the need for regular
    calibration \cite[authoryears][Westwater2004,Kadygrov2015]. Common
    calibration methods for microwave radiometers are tipping curve calibration
    and the use of an internal reference target. Both methods only require a
    rotatable antenna, can be automatically carried out by the instrument and
    are usually complemented by less frequent manual calibrations with a
    liquid nitrogen target. The calibration procedure of MWRs overall is well
    understood \cite[authoryears][Cimini2004] and does not affect suitabiliy
    for deployment in a long-term unattended mode
    \cite[authoryears][Westwater2004,Cimini2006].

    It is instructive to compare the capabilities of instruments commonly used
    to obtain the same atmospheric variables retrievable from a ground-based
    MWR. Radiosondes are usually treated as a reference measurement of the
    thermodynamic atmospheric state due to their high sensor accuracy and fine
    vertical resolution. However, balloons experience horizontal drift during
    their ascent and need a substantial amout of time to reach the tropopause
    reducing the confidence in the representativity of the measurements for
    a given location. The high cost of sensor and balloon causes radiosonde
    launch sites to be distributed only sparsely and launches typically occur
    only one to four times a day which is to little to appropriately capture
    the daily boundary layer evolution. Aircraft soundings ar available in much
    better temporal resolution but can only be taken where airports exist and
    have similar representativity issues as radiosondes due to the horizontal
    displacement of the aircraft path during the ascent. Remote sensing
    instruments provide volume averaged profiles which have a higher
    representativity of the atmospheric state at the cost of vertical
    resolution \cite[authoryears][Westwater2004]. Good temporal resolution can
    be achieved with remote sensors, depending on the need for temporal
    averaging to reduce instrument noise.  Satellites in a polar orbit have
    global coverage but pass a fixed location only in long intervals and
    temperature and humidity sensors on satellites provide only coarse vertical
    resolution in the boundary layer while being more accurate higher up
    \cite[authoryears][Peckham2000,Sanchez2013]. The resolution of
    a ground-based microwave radiometer however is highest in the boundary
    layer and decreases with altitude \cite[authoryears][Cadeddu2002,Lohnert2004].
    Low-level resolution can additionally be increased by scanning in
    off-zenith directions, if the atmosphere is weakly inhomogeneous in the
    horizontal \cite[authoryears][Cimini2006,Crewell2007,Guldner2013].
    Together with a high achievable temporal resolution this makes ground-based
    MWRs well suited for boundary layer research
    \cite[authoryears][Cimini2006]. The spatial coverage of a network of
    ground-based microwave radiometers obviously cannot compete with the
    coverage of a satellite in a polar orbit but the instruments are mobile,
    inexpensive and relatively compact, allowing a wide range of applications
    \cite[authoryears][Kadygrov2015].

    One might try to counter the lack of vertical information in the boundary
    layer with numerical weather prediction (NWP) models but so far mesoscale
    models alone have not been able to fill this data gap
    \cite[authoryears][Sanchez2013]. In the contrary, the understanding of
    boundary layer processed is essential for the development of
    convective-scale models \cite[authoryears][Martinet2015] but this part of
    the atmosphere is undersampled \cite[authoryears][Cimini2015]. Therefore
    current research is concerned with the assimilation of observations from
    ground-based microwave radiometers into NWP models which is currently not
    being done operationally \cite[authoryears][Lohnert2012]. An experiment
    with assimilation of temperature and humidity profiles retrieved from 13
    MWRs by \cite[Cimini2014] resulted in a neutral-to-positive impact on
    forecast skills. Recently a numerical radiative transfer model has been
    developed by \cite[DeAngelis2016] with the goal of direct assimilation of
    radiometer observations.

    Not only numericical weather prediction benefits from additional
    information about the vertical state of the atmosphere, but also other
    forms forecasts. Most prominent are short-term severe weather forecasts
    which stongly depend on accurate vertical profiles of the atmosphere
    \cite[authoryears][Lohnert2012]. \cite[Madhulatha2013,Cimini2015] used
    radiometer observations to derive indices of severe convective activity and
    fog. \cite[Chan2013] analysed the flucutations of brightness temperatures
    measured by a MWR to assess low-level wind shear, crucial for the safe
    operation of an airport. Aside from airports, such information are also
    valuable for the safety nuclear power plants or the assessment of local
    pollution propagation \cite[authoryears][Westwater1972,Kadygrov2015].
    
    To conclude this section, a few notable publication shall be mentioned.
    A review of remote sensing by ground-based sensors in the microwave and
    millimeter-wave region was given by \cite[Westwater2004]. Aside from the
    physical basics and aspects of instrument construction, the authors also
    discuss derivable thermodynamic quantities and give an overview of
    retrieval techniques. A extensive and general review of remote sensing of
    the lower troposphere including a section about radiometers has recently
    been given by \cite[Wulfmeyer2015]. Retrieval techniques are the focus of
    a currently unpublished article by \cite[Turner2013] which can be obtained
    from \hyphenatedurl{http://www.nssl.noaa.gov/users/dturner/public_html/metr5970/retrieval_uncertainty_paper.v4.pdf}.
    The authors have additionally included example applications and discuss
    recent developments.

\stopsection


\startsection[title={Prior Work in Innsbruck}]

    The Institute of Atmospheric and Cryospheric Sciences at the University of
    Innsbruck operates a passive microwave radiometer as part of the Innsbruck
    Box project, whose objective it is to study boundary layer dynamics in
    complex terrain. Consequently there have been efforts to retrieve
    temperature and humidity profiles from brightness temperature measurements
    of the radiometer. Research in Innsbruck has been carried out by
    \cite[Massaro2013] in his master's thesis and a subseqent publication
    \cite[authoryears][Massaro2015], followed by a contribution earlier this
    year by \cite[Meyer2016].
    
    Retrievals for the Innsbruck site have been done so far by linear
    regression techniques with a focus on retrieval performance of profiles
    with a surface-based or elevated temperature inversion and the inclusion of
    regressors other than brightness temperatures in the retrieval process. The
    desire to include additional information stems from Innsbruck's location in
    an Alpine valley at an altitude of approximately 600 m, surrounded by
    mountains of over 2000 m height. This complex terrain provides
    a challenging environment for radiometer retrievals as the boundary layer
    evolution is strongly affected by the topography. But the location in the
    mountains also provides data not available in flat regions: weather
    stations exist at different heights in the mountains, providing in-situ
    profiles of the inner-Alpine atmosphere. Even though these measurements
    are affected by surface layer effects they contain valuable information
    about the state of the lower troposphere. Regression techniques allow easy
    integration of such information into a retrieval. The relationship between
    the valley atmosphere measured by the radiometer and surface layer
    measurements obtained from weather stations does not have to be explicitly
    known but is automatically inferred by the regression.

    \cite[Massaro2015] found that the linear regression retrieval performance
    in Innsbruck is comparable performance to flat regions. They found
    significant improvement in the accuracy of the temperature profile at
    heights where additional information are included and for specialized
    retrievals, which categorize the atmospheric state e.g. by stability. Their
    biggest problem remained the accurate retrieval of elevated temperature
    inversions. \cite[Meyer2016] expanded on these results, taking into account
    a greater number of surface observations from various heights and distances
    from the radiometer. The inclusion of data from these stations combined
    with specialized retrievals allowed even the retrieval of elevated
    temperature inversions with satisfying accuracy.

\stopsection


\startsection[title={Objectives and Outline}]

    In this thesis the retrieval problem is approached from a Bayesian
    perspective. One of the major benefits of the Bayesian approach is the
    assessment of uncertainty of a retrieved profile. This applies to
    retrievals using linear regression but more importantly allows the
    integration of a radiative transfer model into the retrieval yielding
    the well established optimal estimation technique aka 1D-VAR.

    Considering the good results of the prior work in Innsbruck using linear
    regression, there seems to be little incentive to fundamentally change the
    retrieval method. But there are aspects of regression techniques that are
    disadvantageous. Since regression is a purely statistical method any
    knowledge of the underlying physics is - if at all - only implicitly
    included. Introducing additional regressors such as surface temperature
    measurements provide no hard constraints for the retrieval. Similarly
    there is no explicit connection between adjacent atmospheric layers because
    in training the problem decouples into individual regressions for each
    layer. Physically reasonable behavior of retrieved profiles (as usually
    observed in practice) is a consequence of a representative climatology.
    But there is always the possibility of a situation appearing that is far
    away from the climatological range leading to poor regression performance.
    Innsbruck for example has the problem that radiosondes are only launched at
    night, which likely leads to worse retrieval performance during the day
    (without a reasonably sized test dataset of daytime radiosondes this is not
    easy to verify).  Other locations might not even have a training set of
    radiosonde profiles available for training. Model forecasts can be used for
    such locations (CITE THIS) but these are affected by problems associated
    with NWP.

    Optimal estimation techniques following a Bayesian approach improve on
    these shortcomings of regression. Uncertainty assessment is natural to all
    Bayesian methods since entire distributions are propagated through the
    retrieval instead of just values. Physical knowledge, e.g. from a radiative
    transfer model, is explicitly incorporated during retrieval making it
    possible to accurately obtain profiles in situations not covered by
    a climatology. Adding additional instruments to the retrieval scheme seems
    mathematically more elegant (EHHH) than in regression methods and is
    generally more flexible (CITE IPT PAPER?). These benefits come at a cost.
    The radiative transfer model only works with profiles of temperature,
    humidity and cloud liquid water making retrievals of individual variables
    impossible. Computational costs are higher, although this concern is
    becoming increasingly irrelevant with faster models and computers. Even
    though optimal estimation methods do not require a climatology, the data
    that is used must have well characterized errors and known systematic
    biases. This makes the setup much more complicated than that of
    a regression method.

    The goal of this thesis is to provide such a setup for profile retrievals
    in Innsbruck paving the way to operational retrievals useful for boundary
    layer research and weather forecasting. It is an exploratory study as
    variational techniques have never been used for retrievals in Innsbruck.
    Most of the theoretical background of the method and radiative transfer
    is shown here in order to build a deep understanding of the method and all
    its assumptions and implications. Results of a first retrieval
    implementation are shown and compared to regression retrievals. Finally,
    shortcomings of the presented work and ideas for future research are
    given.

    TODO: Although not all papers deal with temperature/humidity and some use
    infrared radiation instead of microwaves, the methods of retrieval are the
    same and all papers are considered here when discussing methodology.
    Since the principles and techniques for all are the same, references are
    mixed and include liquid retrievals as well.

\stopsection

